Towards a miniKanren with fair search strategies

### 2019-06-20 Paper/paper.tex
quasi-camera-ready with line number
updates from Dan

\documentclass[format=acmlarge, review=false, authordraft=false]{acmart}

@@

\title{Towards a miniKanren with fair search strategies}
\author{Kuang-Chen Lu}
\affiliation{Indiana University}
\author{Weixi Ma}
\affiliation{Indiana University}
\author{Daniel P. Friedman}
\affiliation{Indiana University}

@@

\maketitle

@@

miniKanren is a family of relational programming languages.
\citet{Friedman:2005:RS:1121583,friedman_reasoned_2018} introduce
miniKanren and its implementation in \emph{The Reasoned Schemer} 
and \emph{The Reasoned Schemer, 2nd Ed} (TRS2). After \mbox{\citet{Hemann_2016}}
describe microKanren, a minimal core of miniKanren comprised of only 54 LOC, 
miniKanren has been implemented in many other languages, including
multiple ones using the same language (e.g. OCanren\citep{kosarev2018typed}). 
% \citet{byrd2017unified} have demonstrated that miniKanren problems that at 
% first reading seem impossible are indeed possible by using a Scheme interpreter 
% written in miniKanren, thereby having access to logic variables within Scheme.
As demonstrated in \citet{byrd2017unified}, miniKanren can be used to
naturally express difficult computations, such as using an interpreter
to perform example-based program synthesis, or using a proof checker
as a theorem prover.
% miniKanren.org contains links to many miniKanren implementations and the seeds 
% of many unusual problems and their solutions in miniKanren.
The papers, talks, and tutorials on \url{miniKanren.org} present many
other unusual problems, and their solutions in miniKanren.

@@

Our research compares four search strategies with different features in fairness
(\autoref{table-fairness}). To summarize our contributions, we
\begin{itemize}
	\item propose and implement \textbf{b}alanced \textbf{i}nterleaving 
	depth-first search (\DFSbi{})
	\item propose and implement \textbf{f}air depth-first search (\DFSf{})
	\item implement in a new way the standard breath-first search (\BFS) by 
	\citet{seres1999algebra}
	We refer to \BFSser{} as their implementation and \BFSimp{} as our new 
	one. We formally prove that the two implementations are 
	semantically equivalent, however, \BFSimp{} runs faster in all benchmarks 
	and 
	is shorter.
\end{itemize}

\begin{table}[h]
	\begin{tabular}{|c|c|c|}
		\hline 
		Search Strategies & disj & conj \\ 
		\hline 
		\DFSi & unfair & unfair \\ 
		\hline 
		\DFSbi & almost-fair & unfair \\ 
		\hline 
		\DFSf & fair & unfair \\ 
		\hline 
		\BFS & fair & fair \\ 
		\hline 
	\end{tabular} 
	\caption{Fairness of all search strategies}
	\label{table-fairness}
\end{table}

@@

Now we elaborate fairness by running more queries about \repeato{}. We never 
use \runstar{} in this paper because fairness is more interesting when we 
have an unbounded number of answers. It is perfectly fine, however, to use 
\runstar{} with any search strategy.

@@

In this section, we review the implementation of interleaving depth-first 
search (\DFSi). We focus on parts that are relevant to this paper. TRS2,
chapter 10 and the appendix, ``Connecting the wires'', 
provide a comprehensive description of the miniKanren 
implementation but limited to unification constraints ($\equiv$).
Fig.~\ref{DFSi-0} and Fig.~\ref{DFSi-1} show parts that are later compared 
with other search strategies. We follow some conventions to name variables: 
\texttt{s}s name states; \texttt{g}s (possibly with subscript) name goals; 
variables ending with $^\infty$ name spaces. Fig.~\ref{DFSi-0} shows 
the 
implementation of \disj. The 
first function, \disjtwo, implements binary disjunction. It applies the 
two disjunctive goals to the input state \texttt{s} and composes the two 
resulting spaces with \appendInf{}. The following syntax 
definitions say \disj{} is right-associative. Fig.~\ref{DFSi-1} 
shows the implementation of \conj. The first function, \conjtwo{}, 
implements binary conjunction. 
It applies the \emph{first} goal to the input state, then applies the second 
goal to states in the resulting space. The helper function 
\appendMapInf{} applies its input goal to states 
in its input space and composes the resulting spaces. It reuses 
\appendInf{} for space composition. The following syntax 
definitions say \conj{} is also right-associative.

@@

\section{related works}

In this section, we describe related works. 
\citet{yang2010adventures} points 
out that a disjunct complex would be `fair' if it were a full and balanced tree.
\citet{seres1999algebra} describe a breadth-first search 
strategy. We present another implementation. Our implementation is semantically 
equivalent to theirs. But, ours is shorter and performs better in comparison 
with a straightforward translation of their Haskell code.
\citet{rozplokhas2018improving} address the non-commutativity of conjunction, 
while our work about \disj{} fairness addresses the non-commutativity of 
disjunction.

@@

We devise two new search strategies (i.e., balanced interleaving DFS 
(\DFSbi{}) and fair DFS (\DFSf{})) and devise a new 
implementation of \BFS{} (\BFSimp{}). These strategies have different features 
in fairness: \DFSbi{} has an almost-fair \disj{} and unfair \conj{}. 
\DFSf{} has fair \disj{} and unfair \conj{}. \BFS{} has both fair
\disj{} and fair \conj{}. No search strategy in this paper combines unfair 
\disj{} and fair \conj{}. This is because we haven't seen a case where this 
kind of search strategies would be interesting.

>>>>

\documentclass[acmlarge, review=true]{acmart}
%TODO 1. delete `, review=true' ; 2. delete this line .
\usepackage[
    type={CC},       
    modifier={by-sa},
    version={4.0},   
]{doclicense}

\settopmatter{printacmref=false}
\fancyfoot{}

\makeatletter
\def\@formatdoi#1{}
\def\@permissionCodeOne{miniKanren.org/workshop}
\def\@copyrightpermission{\doclicenseThis} % your choice of text
\def\@copyrightowner{Copyright held by the author(s).} % your choice
\makeatother

\copyrightyear{2019}
\setcopyright{rightsretained}

\acmMonth{8}
\acmArticle{1}

@@

\title{Towards a miniKanren with fair search strategies}
\author{Kuang-Chen Lu}
\affiliation{%
  \institution{Indiana University}
  \country{USA}}
\email{kl13@iu.edu}
\author{Weixi Ma}
\affiliation{%
  \institution{Indiana University}
  \country{USA}}
\email{mvc@iu.edu}
\author{Daniel P. Friedman}
\affiliation{%
  \institution{Indiana University}
  \country{USA}}
\email{dfried@indiana.edu}

@@

\maketitle
%\thispagestyle{empty}
%TODO uncomment the line above to be camera-ready

@@

miniKanren is a family of relational programming languages.
\citet{Friedman:2005:RS:1121583,friedman_reasoned_2018} introduce
miniKanren and its implementation in \emph{The Reasoned Schemer} 
and \emph{The Reasoned Schemer, 2nd Ed} (TRS2). \mbox{\citet{Hemann_2016}}
describe microKanren, a minimal core of miniKanren comprised of only 54 LOC, 
miniKanren has been implemented in many other languages, including
multiple ones using the same language (e.g. OCanren\citep{kosarev2018typed}). 
% \citet{byrd2017unified} have demonstrated that miniKanren problems that at 
% first reading seem impossible are indeed possible by using a Scheme interpreter 
% written in miniKanren, thereby having access to logic variables within Scheme.
As demonstrated in \citet{byrd2017unified}, miniKanren can be used to
naturally express difficult computations, such as using an interpreter
to perform example-based program synthesis, or using a proof checker
as a theorem prover.
% miniKanren.org contains links to many miniKanren implementations and the seeds 
% of many unusual problems and their solutions in miniKanren.
The papers, talks, and tutorials on \url{miniKanren.org} present many
other unusual problems, and their solutions in miniKanren.

@@

Our research compares four search strategies with different features in fairness
(\autoref{table-fairness}). To summarize our contributions, we
\begin{itemize}
	\item propose and implement \textbf{b}alanced \textbf{i}nterleaving 
	depth-first search (\DFSbi{})
	\item propose and implement \textbf{f}air depth-first search (\DFSf{})
	\item implement in a new way the standard breath-first search. 
	We refer to \BFSser{} as the original implementation by 
	\citet{seres1999algebra} and \BFSimp{} as our new 
	one. We formally prove that the two implementations are 
	semantically equivalent, however, \BFSimp{} runs faster in all benchmarks 
	and 
	is shorter.
\end{itemize}

\begin{table}[h]
	\begin{tabular}{|c|c|c|}
		\hline 
		Search Strategies & disj & conj \\ 
		\hline 
		\DFSi & unfair & unfair \\ 
		\hline 
		\DFSbi & almost-fair & unfair \\ 
		\hline 
		\DFSf & fair & unfair \\ 
		\hline 
		\BFSimp, \BFSser & fair & fair \\ 
		\hline 
	\end{tabular} 
	\caption{Fairness of all search strategies}
	\label{table-fairness}
\end{table}

@@

Now we elaborate fairness by running more queries about \repeato{}. We never 
use \runstar{} here because fairness is more interesting when we 
ask a bounded number of answers. It is perfectly fine, however, to use 
\runstar{} with any search strategy.

@@

In this section, we review the implementation of interleaving depth-first 
search (\DFSi). We focus on parts that are relevant to other strategies. TRS2,
chapter 10 and the appendix, ``Connecting the wires'', 
provide a comprehensive description of the miniKanren 
implementation but limited to unification constraints ($\equiv$).
Fig.~\ref{DFSi-0} and Fig.~\ref{DFSi-1} show parts that are later compared 
with other search strategies. We follow some conventions to name variables: 
\texttt{s}s name states; \texttt{g}s (possibly with subscript) name goals; 
variables ending with $^\infty$ name spaces. Fig.~\ref{DFSi-0} shows 
the 
implementation of \disj. The 
first function, \disjtwo, implements binary disjunction. It applies the 
two disjunctive goals to the input state \texttt{s} and composes the two 
resulting spaces with \appendInf{}. The following syntax 
definitions say \disj{} is right-associative. Fig.~\ref{DFSi-1} 
shows the implementation of \conj. The first function, \conjtwo{}, 
implements binary conjunction. 
It applies the \emph{first} goal to the input state, then applies the second 
goal to states in the resulting space. The helper function 
\appendMapInf{} applies its input goal to states 
in its input space and composes the resulting spaces. It reuses 
\appendInf{} for space composition. The following syntax 
definitions say \conj{} is also right-associative.

@@

\section{related research}

In this section, we describe related research. 
\citet{yang2010adventures} points 
out that a disjunct complex would be `fair' if it were a full and balanced tree.
\citet{seres1999algebra} describe a breadth-first search 
strategy. We present another implementation. Our implementation is semantically 
equivalent to theirs. But, ours is shorter and performs better in comparison 
with a straightforward translation of their Haskell code.
\citet{rozplokhas2018improving} address the non-commutativity of conjunction, 
while our work about \disj{} fairness addresses the non-commutativity of 
disjunction.

@@

We devise two new search strategies (i.e., balanced interleaving DFS 
(\DFSbi{}) and fair DFS (\DFSf{})) and devise a new 
implementation of \BFS{} (\BFSimp{}). These strategies have different features 
in fairness: \DFSbi{} has an almost-fair \disj{} and unfair \conj{}. 
\DFSf{} has fair \disj{} and unfair \conj{}. \BFS{} has both fair
\disj{} and fair \conj{}. No search strategy here combines unfair 
\disj{} and fair \conj{}. This is because we haven't seen a case where this 
kind of search strategies would be interesting.
